Viscous Transport in Eroding Porous Media

### 2019-08-21 docs/transport/erosion_transport.tex
small changes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
\begin{itemize}
  \item Use novel near-singular Barycentric integration scheme for
    Stokes equation outlined in~\cite{bar-wu-vee2015}. (Remember to
    include historically significant reference Shelley
    \cite{bak-she1986}).
  \item Need to compute one additional derivative since we need the
    derivative of the velocity to find the vorticity and stress tensor
  \item Annealing algorithm to create initial body configurations of
  \item Tracer trajectories at different time steps
  \item Compute the anomalous dispersion rate using similar techniques
  as the paper with Pietro~\cite{dea-qua-bir-jua2018}
  \item Need a novel way to do time stepping through the complex
    geometry
\end{itemize}

\paragraph{Limitations}
\begin{itemize}
  \item Bodies are still pinned
  \item Bodies are still two dimensional
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by defining the main variables used to model erosion.  A more
detailed description is in our previous work~\cite{qua-moo2018}.  We
consider flows inside a confined geometry $\Omega$ that contains $M$
eroding bodies with boundary $\gamma_\ell$.  The boundary of the
bounding geometry is $\Gamma$, and the fluid domain is bounded by $\bd
\Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$.  On each
eroding body, $\nn$ is the normal vector pointing into the body, and
$\ss$ is the unit tangent vector pointing in the counterclockwise
direction.  The normal vector on $\Gamma$ is the outward unit normal.
Neglecting inertial forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}\\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass} \\
    \uu = 0, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the bodies} \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity} \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$.

@@

To quantify the tortuosity, we compute the relative velocity
(Figure~\ref{fig:Eroding100tort}(a)) and the local tortuosity
(Figure~\ref{fig:Eroding100tort}(b)) of 1000 tracers placed immediately
to the left of the porous region at $x=-1$.  The porosity at this
instance in time is $62.98\%$ and the maximum velocity at $x=-1$ is
$3.90 \times 10^{-4}$. The initial velocity of the tracers is
qualitatively similar to the 20 body example
(Figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Compared to
Figure~\ref{fig:Eroding20tort}(b), the local tortuosity is much more
discontinuous.  These discontinuities can be explained by examining the
trajectories of tracers in Figure~\ref{fig:Eroding100tort}(c).  Here,
there are many instances of neighboring trajectories that are deflected
apart from one another as they tend to a stagnation point in the flow,
and this results in trajectories with significantly different lengths.
At this instance in time one the tracers travels 25.5\% farther than it
would have if the bodies had been absent, and the average tracer
travelled ???\% farther than it would have if the bodies were absent.
Therefore, at this instance in time, the tortuosity is $1.???$.
\todo[inline]{S-H: Fill in the numbers in this paragraph}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Contributions}
Our first main contribution is extending the work of Barnett et
al.~\cite{bar-wu-vee2015} so that high-order accurate derivatives of
solutions of the Stokes equations can be computed.  Using this new
quadrature method, we can accurately compute erosion rates, and
therefore stably simulate erosion.  Moreover, we use the quadrature
method to visualize the vorticity in the fluid bulk.  The vorticity is
convenient to visualize since, in the zero Reynolds number limit, it is
equivalent to the shear stress on the eroding bodies, and therefore its
magnitude is the erosion rate.

Our second contribution is computing trajectories in the eroded
geometries.  These trajectories are used to characterize the transport
by computing local tortuosities of the trajectories, the tortuosity of
the eroded geometry, and the anomalous dispersion rates in the geometry.
Again, by using an appropriate quadrature method~\cite{bar-wu-vee2015},
and a fourth-order Runge-Kutta method, we are able to stably compute
trajectories that spend large amounts of time in the slow velocity
regions near the eroding bodies, and computing these trajectories
reliably is essential to characterize the flow.  We compare our
tortuosity and dispersion rates with models available in the literature.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Limitations}
As in our previous work~\cite{qua-moo2018}, this work considers
two-dimensional immobile bodies. However, this new work does not have a
limitation on the packing of the bodies. Compared to our previous work,
we are able to simulate much denser suspensions with a comparable number
of discretization points.  While we do investigate transport in this
work, we only consider passive particles that are undergoing only
advection. To model other physical processes, such as contaminant
transport, we plan to simulate a finite Peclet number
advection-diffusion equation in eroded geometries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations}
\label{sec:formulation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\cite{bak-she1986}

We start by defining the main variables used to model erosion.  A more
detailed description is in our previous work~\cite{qua-moo2018}.  We
consider flows inside a confined geometry $\Omega$ that contains $M$
eroding bodies with boundary $\gamma_\ell$.  The boundary of the
bounding geometry is $\Gamma$, and the fluid domain is bounded by $\bd
\Omega = \Gamma \cup \gamma_1 \cup \cdots \cup \gamma_M$.  On each
eroding body, $\nn$ is the normal vector pointing into the body, and
$\ss$ is the unit tangent vector pointing in the counterclockwise
direction.  The normal vector on $\Gamma$ is the outward unit normal.
Neglecting inertial forces, the governing equations are
\begin{equation}
\label{eqn:erosionModel}
  \begin{split}
    \mu \Delta \uu = \grad p, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of momentum}\\
    \grad \cdot \uu = 0, &\hspace{20pt} \xx \in \Omega, \gap 
      &&\mbox{\em conservation of mass} \\
    \uu = 0, &\hspace{20pt} \xx \in \gamma, \gap 
      &&\mbox{\em no slip on the bodies} \\
    \uu = \UU, &\hspace{20pt} \xx \in \Gamma, \gap 
      &&\mbox{\em outer wall velocity} \\
    \Vn = \CE \, \abs{\tau}, &\hspace{20pt} \xx \in \gamma,
      &&\mbox{\em erosion model}.
  \end{split}
\end{equation}
Here, $\uu$ is the fluid velocity, $p$ is the pressure, $\UU$ is a
prescribed Hagen-Poiseuille velocity field, $\Vn$ is the normal velocity
of $\gamma$, and
\begin{align}
  \tau = -(\nabla \uu + \nabla \uu^T) \nn \cdot \ss
  \label{eqn:shearStress}
\end{align}
is the shear stress on $\gamma$.

@@

To quantify the tortuosity, we compute the relative velocity
(Figure~\ref{fig:Eroding100tort}(a)) and the local tortuosity
(Figure~\ref{fig:Eroding100tort}(b)) of 1000 tracers placed immediately
to the left of the porous region at $x=-1$.  The porosity at this
instance in time is $62.98\%$ and the maximum velocity at $x=-1$ is
$3.90 \times 10^{-4}$. The initial velocity of the tracers is
qualitatively similar to the 20 body example
(Figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Compared to
Figure~\ref{fig:Eroding20tort}(b), the local tortuosity is much more
discontinuous.  These discontinuities can be explained by examining the
trajectories of tracers in Figure~\ref{fig:Eroding100tort}(c).  Here,
there are many instances of neighboring trajectories that are deflected
apart from one another as they tend to a stagnation point in the flow,
and this results in trajectories with significantly different lengths.
At this instance in time one the tracers travels 25.5\% farther than it
would have if the bodies had been absent, and the average tracer
travelled 12\% farther than it would have if the bodies were absent.
Therefore, at this instance in time, the tortuosity is $1.12$.
